---
title: "W2_proj"
author: "SIYI"
date: "10/19/2020"
output: html_document
---

## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv")
dim(activity) #17568 rows and 3 columns
names(activity)
length(unique(activity$date)) #61 unique dates.
```

## What is mean total number of steps taken per day?

```{r}
#total number of steps taken per day
total_number_steps <- with(activity, tapply(steps, as.factor(activity$date), sum))
summary(total_number_steps)
```
Without removing the missingness, the mean total number of steps taken per day is 10766, median is 10765.

```{r}
hist(x = total_number_steps, main = "Total number of steps taken per day", xlab = "Total number of steps")
```


## What is the average daily activity pattern?

```{r}
activity1 <- activity[!is.na(activity$steps),]
mean_steps <- with(activity1, tapply(steps, activity1$interval, mean))
interval <- levels(as.factor(activity1$interval))
plot(interval, mean_steps, type ="l", main = "Average number of steps taken", xlab = "interval", ylab = "Mean steps")
```

```{r}
dt <- data.frame(mean_steps, interval)
dt[dt$mean_steps==max(dt$mean_steps),]
```
Interval 835 contains the max number of steps.s

## Imputing missing values

```{r check the missingness}
n_na <- activity[(which(is.na(activity))),] #2304 records have missing values
na_steps <- activity[(which(is.na(activity$steps))),] #2304
na_date <- activity[(which(is.na(activity$date))),] #0
na_interval <- activity[(which(is.na(activity$interval))),] #2304
```

Only `steps` has misssingness. The total number of missingness is 2304.

```{r}
#replace NA with mean
n_na$steps <- mean_steps
complete_act <- rbind(n_na, activity1)
```

```{r}
total_number_steps2 <- with(complete_act, tapply(steps, as.factor(complete_act$date), sum))
#Make a histogram 
hist(total_number_steps2, main = "Total number of steps taken per day", xlab = "Total number of steps")
```

```{r}
summary(total_number_steps2)
```
The mean total number of steps taken per day is 10766, median is 10766.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
complete_act$days <- weekdays(as.Date(complete_act$date))
# find weekend features in the dataset
weekend_feature <- grep("Saturday|Sunday", complete_act$days, ignore.case = T)
# subset data of the weekend
weekend_dt<- complete_act[weekend_feature, ]
weekend_dt$weekday <- "weekend"

# subset data of the weekday
weekday_dt <- subset(complete_act,complete_act$days!=weekend_feature)
weekday_dt$weekday <- "weekday"
new <- rbind(weekday_dt, weekend_dt)
```

```{r}
library(ggplot2)
mean_number_steps <- aggregate(steps~ interval+weekday, new, mean)
ggplot(mean_number_steps, aes(interval, steps)) +
    geom_line(stat = "identity", aes(color = weekday)) +
    theme_gray() +
    facet_grid(weekday ~ ., scales="fixed", space="fixed") +
    labs(x="Interval", y=expression("Number of Steps")) +
    ggtitle("Steps Per Interval")
```

